### 2020-01-13 paper/sex_selection_spacing-ver3.tex
Add notes on prediction set-ups

One of the effects of the increasing use of sex selection over time is that the total 
fertility rate will tend to overestimate the decline in fertility.
``Tempo'' effects, such as those that arise from postpone of childbearing, are studied 
extensively in the literature \citep{Bongaarts1999}.
With development, women tend to have their first child later as a result of increasing 
years spent in school and later marriage, but, as discussed above, spacing between births 
also tend to shorten as women become more likely to join the labor market, which will 
tend to counter the downward effect on the TFR.

India has also seen an increasing age of marriage, but birth spacing has also become
much longer as women leave the labor market and there is substantial use of 
sex selection.
The result is that the TFR becomes an increasingly biased measure of cohort fertility
\citep{Ni-Bhrolchain2011}.

@@

>>>>

One of the effects of the increasing use of sex selection is that the total fertility 
rate will tend to overestimate the decline in fertility.
``Tempo'' effects, such as those that arise from postpone of childbearing, are studied 
extensively in the literature \citep{Bongaarts1999}.
With development, women tend to have their first child later as a result of increasing 
years spent in school and later marriage, but, as discussed above, spacing between births 
also tend to shorten as women become more likely to join the labor market, which will 
tend to counter the downward effect on the TFR.

India has also seen an increasing age of marriage, but birth spacing has also become
much longer as women have become less likely to join the labor market and with the 
substantial use of sex selection.
The result is that the TFR may have become an increasingly biased measure of cohort fertility.%
\footnote{
I use ``bias'' because what I am interested in is the cohort fertility \citep{Ni-Bhrolchain2011}
}

@@

I calculate TFR by urban/rural and the education groups used above.
The main difference from the standard TFR measure is that I only use births up to parity
four.
Hence, the TFRs are not directly comparable those in the NFHS reports, although the
differences are smaller the higher the education group and the most recent the survey
as fertility has fallen over time.
Otherwise, I follow the procedure used in the NFHS reports and count the number of births
in months 1 to 36 before the survey month (births in the survey month are not included)
for five-year age groups.

Because NFHS-1 and NFHS-2 did not interview all women between 15 and 49 about fertility
the number of women is based on the household rosters, assuming that never-married 
women have had no children.

To predict cohort fertility based on the models above, I estimate for each spell the 
probability of progression to the next parity.%
\footnote{
This includes the spell from marriage to first birth, even those results are not 
examined above because of the problems with capturing the exact timing of marriages
for the first two rounds of the survey.
The estimated progression probabilities should, however, be less affected by this
problem.
}

To make the predictions depend solely on sample composition and the estimate, I
use the age at marriage for each woman and predict the likelihood of progressing to
each parity, assuming two years increases in age between each parity.
For example, for a woman married at age 20, I use age 22 for predicting progression 
to the second birth, 24 for the third, and 26 for the fourth.

The parity progression depends on the sex composition for prior children for
parity two and above.
For example, the likelihood of a second birth is the 
probability of having a first birth multiplied with
sum of the 
probability of having a son as the first child conditional on having a first child 
multiplied with parity progression probability if the first child is a son 
and the 
probability of having a girl as the first child  conditional on having a first child
multiplied with the parity progression probability if the first child is a girl.
$$
prob(2nd) = prob(1st) * (prob(son|1st)*prob(2nd|son) + prob(girl|1st) * prob(2nd|girl)
$$

The survey rounds do not coincide entirely with the periods I use for the model.
For the NFHS-4, I use the model results for the 2005--2016 period, 
for NFHS-3 the 1995-2004, NFHS-2 1985-1994, and for NFHS-1 1972-1984.
